Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-07-24 paper/manuscript.tex

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[1*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[2*]{John D. Chodera}
\author[1,10]{Maximilian Kuhn}
\author[2]{Hannah E. Bruce Macdonald}
\author[1*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[11]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[2,8]{Andrea Rizzi}
\author[1]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[9]{Gary Tresadern}
\author[7]{Huafeng Xu}
\affil[1]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[2]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[3]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, USA}
\affil[4]{National Institutes of Health, Bethesda, MD, USA}
\affil[6]{University of Colorado Boulder, Boulder, CO, USA}
\affil[7]{Silicon Therapeutics, Boston, MA, USA}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[9]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340,Belgium}
\affil[10]{Cresset, Cambridgeshire, UK}
\affil[11]{Molecular Sciences Software Institute, Blacksburg VA, USA}
\corr{john.chodera@choderalab.org}{JDC}
\corr{dmobley@mobleylab.org}{DLM}
\corr{antonia.mey@ed.ac.uk}{ASJSM}
\corr{mail@julienmichel.net}{JM}
\corr{michael.shirts@colorado.edu}{MRS}
\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Month, Day, Year}
\dateaccepted{Month, Day, Year}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that cannot exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the efficient computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly. 

@@

In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
%\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}
\end{frontmatter}
\todototoc
\listoftodos
%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%
\todo[inline]{}The current Overleaf document is \url{https://www.overleaf.com/5136551543vgjqbkzpmqkp}
\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor, the transfer of a small molecule from an aqueous to apolar phase~\cite{zwanzig1954hightemperature}, or effects of protein side chain mutations on e.g. binding affinities. 
These calculations use non-physical\footnote{Here, the non-physical nature of the transformation is referred to as "alchemical", a term coined by Tembre and McCammon in Ref.~\cite{tembre1984ligandreceptor}.} intermediate states in which the chemical identity of some portion of the system (such as a small molecule ligand or receptor sidechain) is changed by modifying the potential governing the interaction of the atoms being modified, inserted, or deleted with their environment. 
Fig.~\ref{fig:fig_what_is_alchemy} illustrates common free energy changes that may be difficult to compute with unbiased molecular dynamics methods, but are more readily tractable with alchemical methods.
In alchemical simulations, the introduction of intermediate \textit{alchemical states} that bridge the high-probability regions of configuration space between the two physical endstates of interest permits the free energy of large transformations to be robustly computed.
Alchemical calculations can be used in a variety of scenarios, such as: 
\item computing the free energy of a conformational change for a molecule with a high barrier to interconversion (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{A});
\item computing partition ($\log P$) or distribution ($\log D$) coefficients between environments (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{B})~\cite{rustenburg2016measuring, bosisio2016blinded} 
\item or between compartments or membranes (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{C})~\cite{corey2019insights}. 

Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: 
\begin{itemize}
\item a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{D})\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; 
\item the entire ligand can be alchemically transferred from protein to solvent in an absolute binding free energy calculation (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{E})~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; 
\item and small alchemical modifications can be made between chemically related ligands to estimate relative differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{F})~\cite{wang2015accurate,mey2016blinded,song2019using,gapsys2020large,kuhn2020automated}.
\end{itemize}
%

Once an alchemical simulation procedure is performed (which usually involves more than one simulation), the data must be analyzed to estimate the free energy for the transformation of interest.
Early work used simple but statistically suboptimal estimators for this: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or thermodynamic integration (TI), for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamicsa, merz1989free}. %
More recent developments have seen new, highly efficient statistical estimators that make better use of all the data, often building on the Bennett acceptance ratio (BAR)~\cite{bennett1976efficient}, producing multistate generalizations~\cite{shirts2008statisticallya} or removing the need for global equilibrium~\cite{wu2016multiensemble, mey2014xtram, wu2014statistically}.

Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{shirts2003extremely,shirts2005solvation,vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. 
This foundational work, combined with the methodological, technological, and hardware improvements of the last 5--10 years, have led to an explosion of interest and direct commercial application of the technologies~\cite{wang2015accurate, fratev2019improved, schindler2020largescale, cournia2017relative, sherborne2016collaborating}.
%

As the field of molecular simulation moves towards millisecond timescales, accurate alchemical calculations on even more challenging problems will become reasonable to perform. In the meantime, today's users may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This Best Practice guide provides current recommendations and tips for users of all experience.  
%
\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} \textbf{A}: Change in free energy due to a conformational change of the molecule across a high barrier. \textbf{B}: Partition coefficient such as $\log P$ or $\log D$ depend on a change in free energy between different phases; here, as an example the partition coefficient between methanol and water is shown. \textbf{C}: Free energy difference associated with the insertion of a molecule into a membrane. \textbf{D}: Effect of mutations of protein or host residues on free energies of binding. \textbf{E}: Absolute free energies of binding of a small molecule to a host (e.g. protein), \textbf{F}: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
What can be expected from this Best Practice guide is a good starting point for new practitioners and a reference for experienced practitioners. For this propose we provide a convenient checklist (Section~\ref{sec:checklist}) to help ensure all calculations meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.
What we assume from a novice practitioner is a moderate experience with molecular simulation packages and concepts. Furthermore, basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association are essential. If you feel unfamiliar with some of these concepts, good starting points can be found in these references~\cite{braun2019best, grossfield2018best, klimovich2015guidelines, shirts2012best}. While reading this Best Practice document, it is important to bare in mind \emph{this is not a review} of all free energy calculation methods at the cutting edge of current research.
Instead the guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
\end{itemize}
%

Some other background information may be needed depending on the nature of the alchemical project. For example, often, if binding poses are not known, docking calculations can be used to generate an initial small molecule binding pose to start alchemical calculations. This will require some basic familiarity on how to perform docking calculations to generate reasonable simulation starting points~\cite{grinter2014challenges}. 
%
As some of the theoretical background can seem daunting, we do, however, provide a guide to the essential theory behind alchemical free energy calculations in Section~\ref{sec:theory}.

In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:prerequisites}), choice and use of correct protocols~(\ref{sec:simulation_protocol_choice}), and finally the best practices that should be used in the analysis of alchemical calculations~(\ref{sec:data_analysis}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).
%

While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future Best Practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item covalent inhibition~\cite{lameira2019predicting}
\item free energies of mutation of protein side chains~\cite{gapsys2016accurate, aldeghi2018accurate}
\item nonspecific binding or multiple binding sites~\cite{gill2018binding}
\item complex association stoichiometry (beyond simple 1:1 association).
\item approximate and often less accurate endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. A number of different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current}. 
\end{itemize}
%

For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    \textit{CPU} -- Central Processing Unit\\
    \textit{BAR} -- Bennett Acceptance Ratio\\
    \textit{FEP} -- Free Energy Perturbation\\
    \textit{GPCR} -- G-Protein Coupled Receptor\\
    \textit{GPU} -- Graphics Processing Unit\\
    \textit{MBAR} -- Multistate Bennett Acceptance Ratio\\
    \textit{MCSS} -- Maximum Common Substructure\\
    \textit{MD} -- Molecular Dynamics\\
    \textit{RMSE} -- Root Mean Square Error\\
    \textit{MUE} -- Mean unsigned error\\
    \textit{SAR} -- Structure-Activity Relationships\\
    \textit{TI} -- Thermodynamic Integration
\end{tcolorbox}

\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
$L$ and $R$ -- generic names for ligand and receptor\\
$K_b^{\circ}$ -- binding  constant \\
$c^{\circ}$ -- standard state concentration \\
$U$ -- potential energy\\
$u$ -- reduced potential describing a thermodynamic state \\
$\Delta G$ -- Gibbs free energy (isothermal isobaric ensemble)\\
$\Delta A$ -- Helmholtz free energy (canonical ensemble)\\
$\Delta f$-- reduced (dimensionless) free energy \\
$\Delta \hat{f}$ -- estimate from an estimator for the reduced free energy\\
$\Gamma$ -- conformation space accessible by simulations \\
$\vec{q}$ -- vector of a single configuration, i.e. $x$, $y$, $z$ coordinates of the simulation system\\
$k_B$ -- Boltzmann constant \\
$Z$ -- partition function \\
$p$ -- pressure \\
$\mu$ -- chemical potential (grand canonical ensemble)\\
$T$ -- temperature \\
$\beta \equiv (k_B T)^{-1}$ -- inverse thermal energy \\
$\vec{\lambda}$ -- alchemical progress parameter \\
$g$ -- statistical inefficiency
\end{tcolorbox}

\begin{tcolorbox}[title=Common terminology, colback=yellow!10!white]
    \emph{ligand} - small molecule binding to a larger host molecule such as a protein (receptor).\\ 
    \emph{lead optimization} - process of designing a drug candidate from an initial lead compound, usually involving optimizing its affinity, possibly selectivity, and various pharmacokinetic parameters.\\
    \emph{docking} -- Method of generate simulation starting structures for ligand receptor systems, when no crystal structure is available\\
    \emph{thermodynamic cycle} - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.\\
    \emph{alchemical} -- states or simulations that contain non-physical intermediates \\
\end{tcolorbox}

@@

\paragraph{Why do alchemical calculations need unphysical intermediate states?}
While it is theoretically possible to estimate the ratio of partition functions from samples collected only at states $\vec{\lambda}_A$ and $\vec{\lambda}_B$, the efficiency of the free energy estimators rapidly decreases as the phase-space overlap between the two states also decreases~\cite{wu2005phasespaceb,wu2005phasespacec}.
Roughly, the phase-space overlap between two thermodynamic states measures the degree to which high-probability conformations (i.e., those with very negative potential energy) in one state are high-probability conformations in the other state as well (see Section~\ref{sec:are-they-good} and Fig.~\ref{fig:fig_what_is_lambda}).

@@

The main difference in these approaches lies instead in the thermodynamic cycle to which this strategy is applied.
For example, a typical thermodynamic cycle for an alchemical absolute binding free energy calculation is represented in Figure~\ref{fig:fig_absolute_thermodynamic_cycle}.
In this case, two independent calculations compute the free energy of removing the interactions between the ligand and its environment in solvent or in the binding site respectively through a series of intermediate states in which the energy terms are only partially deactivated.

@@

\paragraph{Protonation states depend on the pH of the experimental assay}
I think we should have a tiny bit on protonation states. This is something I always easily forget as a non-chemist.

@@

\bibliography{alchemical1,manual}

>>>>

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[1*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[2]{Hannah E. Bruce Macdonald}
\author[2*]{John D. Chodera}
\author[1,10]{Maximilian Kuhn}
\author[1*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[11]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[2,8]{Andrea Rizzi}
\author[1]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[9]{Gary Tresadern}
\author[7]{Huafeng Xu}
%
\affil[1]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[2]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[3]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, USA}
\affil[4]{National Institutes of Health, Bethesda, MD, USA}
\affil[6]{University of Colorado Boulder, Boulder, CO, USA}
\affil[7]{Silicon Therapeutics, Boston, MA, USA}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[9]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340,Belgium}
\affil[10]{Cresset, Cambridgeshire, UK}
\affil[11]{Molecular Sciences Software Institute, Blacksburg VA, USA}
%
\corr{john.chodera@choderalab.org}{JDC}
\corr{dmobley@mobleylab.org}{DLM}
\corr{antonia.mey@ed.ac.uk}{ASJSM}
\corr{mail@julienmichel.net}{JM}
\corr{michael.shirts@colorado.edu}{MRS}
%
\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Month, Day, Year}
\dateaccepted{Month, Day, Year}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that cannot exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the efficient computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly.

@@

In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
%\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}
\end{frontmatter}

\todototoc
\listoftodos

%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%
\todo[inline]{}The current Overleaf document is \url{https://www.overleaf.com/5136551543vgjqbkzpmqkp}
\section{What are alchemical free energy methods?}
\label{sec:intro}
Alchemical free energy calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor, the transfer of a small molecule from an aqueous to apolar phase~\cite{zwanzig1954hightemperature}, or the effects of protein side chain mutations on binding affinities or thermostabilities. 
 These calculations use non-physical\footnote{Here, the non-physical nature of the transformation is referred to as "alchemical", a term coined by Tembre and McCammon in Ref.~\cite{tembre1984ligandreceptor}.} intermediate states in which the chemical identity of some portion of the system (such as a small molecule ligand or protein sidechain) is changed by modifying the potential governing the interactions with the environment for the atoms being modified, inserted, or deleted.

Fig.~\ref{fig:fig_what_is_alchemy} illustrates common free energy changes that may be difficult to compute with unbiased molecular dynamics methods, but are more readily tractable with alchemical methods.
In alchemical simulations, the introduction of intermediate \textit{alchemical states} that bridge the high-probability regions of configuration space between the two physical endstates of interest permits the free energy of large transformations to be robustly computed.
Alchemical calculations can be used in a variety of scenarios, such as: 
\begin{itemize}
\item computing the free energy of a conformational change for a molecule with a high barrier to interconversion (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{A});
\item computing partition ($\log P$) or distribution ($\log D$) coefficients between environments (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{B})~\cite{rustenburg2016measuring, bosisio2016blinded} 
\item or between compartments or membranes (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{C})~\cite{corey2019insights}. 
\end{itemize}

Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: 
\begin{itemize}
\item a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{D})\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; 
\item the entire ligand can be alchemically transferred from protein to solvent in an absolute binding free energy calculation (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{E})~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; 
\item and small alchemical modifications can be made between chemically related ligands to estimate relative differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{F})~\cite{wang2015accurate,mey2016blinded,song2019using,gapsys2020large,kuhn2020automated}.
\end{itemize}

After an alchemical calculation is performed (which generally involves multiple simulations at a variety of alchemical states), the data must be analyzed to compute an estimate the free energy for the transformation of interest.
 Early work used simple but statistically suboptimal estimators for this: free energy perturbation (FEP) used a simple (but highly biased) estimator based on the Zwanzig relation~\cite{zwanzig1954hightemperature} or numerical quadrature via thermodynamic integration (TI), for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamicsa, merz1989free}. %
 More recent developments have seen new, highly efficient statistical estimators that make better use of all the data, often building on the more efficient and less biased Bennett acceptance ratio (BAR)~\cite{bennett1976efficient}, producing multistate generalizations~\cite{shirts2008statisticallya} or removing the need for global equilibrium~\cite{wu2016multiensemble, mey2014xtram, wu2014statistically}.

Subsequent work in the 2000s led to improved implementations of alchemical methods in popular biomolecular simulation packages~\cite{shirts2003extremely,shirts2005solvation,vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. 
 This foundational work, combined with the methodological, technological, and hardware improvements of the last 5--10 years, have led to an explosion of interest and direct commercial application of the technologies~\cite{wang2015accurate, fratev2019improved, schindler2020largescale, cournia2017relative, sherborne2016collaborating}.

As the field of molecular simulation can now routinely access microsecond timescales with the aid of GPUs~\cite{salomon-ferrer2013routine} and millisecond timescales appear to soon be within reach, accurate alchemical calculations on even more challenging problems will become reasonable to perform. 
In the meantime, today's users may find it difficult to get started with these complex calculations whilst also keeping up with the fast pace of change. 
This Best Practices guide provides current recommendations and tips for users of all experience.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} \textbf{A}: Change in free energy due to a conformational change of the molecule across a high barrier. \textbf{B}: Partition coefficient such as $\log P$ or $\log D$ depend on a change in free energy between different phases; here, as an example the partition coefficient between methanol and water is shown. \textbf{C}: Free energy difference associated with the insertion of a molecule into a membrane. \textbf{D}: Effect of mutations of protein or host residues on free energies of binding. \textbf{E}: Absolute free energies of binding of a small molecule to a host (e.g. protein), \textbf{F}: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
What can be expected from this Best Practices guide is a good starting point for new practitioners and a reference for experienced practitioners. 
 For this propose we provide a convenient checklist (Section~\ref{sec:checklist}) to help ensure all calculations comply with currently-understood best practices for alchemical simulation and analysis. 
 Where the best practices are currently unsettled, we highlight areas where further research is needed to identify an unambiguous recommendation.
 It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.

 What we assume from a novice practitioner is a moderate experience with molecular simulation packages and concepts. 
 Furthermore, basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association are essential. If you feel unfamiliar with some of these concepts, good starting points can be found in these references~\cite{braun2019best, grossfield2018best, klimovich2015guidelines, shirts2012best}.

 While reading this Best Practices guide, it is important to bear in mind \emph{this is not a review} of all free energy calculation methods at the cutting edge of current research.
Instead the guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical calculation? 
     \item How do I select an appropriate alchemical protocol? 
     \item What software tools are available to perform alchemical calculations? 
     \item How should I analyze my data and report uncertainties? 
\end{itemize}

Some other background information may be needed depending on the nature of the alchemical project. For example, often, if binding poses are not known, docking calculations can be used to generate an initial small molecule binding pose to start alchemical calculations. This will require some basic familiarity on how to perform docking calculations to generate reasonable simulation starting points~\cite{grinter2014challenges}.

As some of the theoretical background can seem daunting, we do, however, provide a guide to the essential theory behind alchemical free energy calculations in Section~\ref{sec:theory}.

In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:prerequisites}), choice and use of correct protocols~(\ref{sec:simulation_protocol_choice}), and finally the best practices that should be used in the analysis of alchemical calculations~(\ref{sec:data_analysis}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future Best Practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have \emph{not} addressed, along with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item covalent inhibition~\cite{lameira2019predicting}
\item free energies of mutation of protein side chains~\cite{gapsys2016accurate, aldeghi2018accurate}
\item nonspecific binding or multiple binding sites~\cite{gill2018binding}
\item approximate and often less accurate endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. A number of different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current}. 
\end{itemize}

For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    {\bf CPU} -- Central Processing Unit\\
     {\bf BAR} -- Bennett Acceptance Ratio\\
     {\bf FEP} -- Free Energy Perturbation\\
     {\bf GPCR} -- G-Protein Coupled Receptor\\
     {\bf GPU} -- Graphics Processing Unit\\
     {\bf MBAR} -- Multistate Bennett Acceptance Ratio\\
     {\bf MCSS} -- Maximum Common Substructure\\
     {\bf MD} -- Molecular Dynamics\\
     {\bf RMSE} -- Root Mean Square Error\\
     {\bf MUE} -- Mean unsigned error\\
     {\bf SAR} -- Structure-Activity Relationships\\
     {\bf TI} -- Thermodynamic Integration
\end{tcolorbox}

\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
$L$ and $R$ -- generic names for ligand and receptor\\
$K_b^{\circ}$ -- binding  constant \\
$c^{\circ}$ -- standard state concentration \\
$U$ -- potential energy\\
$u$ -- reduced potential describing a thermodynamic state \\
$\Delta G$ -- Gibbs free energy (isothermal isobaric ensemble)\\
$\Delta A$ -- Helmholtz free energy (canonical ensemble)\\
$\Delta f$-- reduced (dimensionless) free energy \\
$\Delta \hat{f}$ -- estimate from an estimator for the reduced free energy\\
$\Gamma$ -- conformation space accessible by simulations \\
$\vec{q}$ -- vector of a single configuration, i.e. $x$, $y$, $z$ coordinates of the simulation system\\
$k_B$ -- Boltzmann constant \\
$Z$ -- partition function \\
$p$ -- pressure \\
$\mu$ -- chemical potential (grand canonical ensemble)\\
$T$ -- temperature \\
$\beta \equiv (k_B T)^{-1}$ -- inverse thermal energy \\
$\vec{\lambda}$ -- alchemical progress parameter \\
$g$ -- statistical inefficiency\\
$\omega$ - overlap matrix\\
$C_t$- discrete-time-normalized fluctuation auto-correlation function\\
$\tau _{eq}$ - integrated auto-correlation time\\
$t_0$ - equilibration time
\end{tcolorbox}

@@

\paragraph{Why do alchemical calculations need unphysical intermediate states?}
While it is theoretically possible to estimate the ratio of partition functions from samples collected only at states $\vec{\lambda}_A$ and $\vec{\lambda}_B$, the efficiency of the free energy estimators rapidly decreases as the phase-space overlap between the two states also decreases~\cite{wu2005phasespaceb,wu2005phasespacec}.
Roughly, the phase-space overlap between two thermodynamic states measures the degree to which high-probability conformations (i.e., those with very negative potential energy) in one state are also high-probability conformations in the other state(see Section~\ref{sec:are-they-good} and Fig.~\ref{fig:fig_what_is_lambda}).

@@

The main difference in these approaches lies instead in the thermodynamic cycle to which this strategy is applied.
For example, a typical thermodynamic cycle for an alchemical absolute binding free energy calculation is represented in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}.
In this case, two independent calculations compute the free energy of removing the interactions between the ligand and its environment in solvent or in the binding site respectively through a series of intermediate states in which the energy terms are only partially deactivated.

@@

\paragraph{Protonation states depend on the pH of the experimental assay}
Care should be taken when preparing ligands and proteins to match the experimental assay conditions in terms of pH, if they are known. The pH of the assay will determine the protonation states of the proteins and ligands. Since the pKa of amino acid residues is known different tools are available for proteins, such as the H++ server, or Maestro are good starting points~\cite{anandakrishnan2012automating, 2020schrodinger}. For ligands often the pKa needs to be deterimined if not known experimentally. There are many different available tools for this purpose, but common choices may be propKa~\cite{olsson2011propka3,sondergaard2011improved}, Chemicalize (\url{https://chemicalize.com/welcome}), or Maestro~\cite{2020schrodinger}. While mostly it can be assumed that protonation states will remain the same for an e.g. ligand protein binding calculation, some care needs to be taken with systems where the protonation state may change upon binding~\cite{onufriev2013protonation}.

@@

\bibliography{alchemical,manual}
